---
title: "ENVS-193DS_homework-03"
format: html
name: "Linnea Lauffer"
date: "2025-05-28"
link: "https://github.com/linnea-lauffer/ENVS-193DS_homework-03.git"
---

# Set Up

```{r Set Up}
#| message: false # use this to make sure messages don't show up
#| warning: false # use this to make sure warnings don't show up
#| toc: true # use this to display a table of contents

# load packages 
library(tidyverse)
library(here)
library(gt)
library(janitor)

my_data <- read_csv(here("data","Personal Data.csv")) # reading in my csv file from the data folder and storing it as an object named my_data
```


# Problem 1. Personal Data

## a. 

I could calculate the mean total hours of sleep for both work nights and non-work nights to compare how my sleep duration varies. This is informative, as I expect that on work nights I may sleep less as I have a night time job and have classes in the late morning, and comparing the average sleep time helps me evaluate how my work schedule affects my rest. 

## b. 

```{r}
my_data_clean <- my_data |> # starting data frame
  clean_names() # clean column names

sleep_summary <- my_data_clean |> # start with cleaned data
  mutate(work_night = as.factor(work_night)) |> # convert work_night column to a factor (categorical variable)
  group_by(work_night) |> # group the data by work_night
  summarise(mean_sleep = mean(total_sleep_hours)) # calculate mean of total_sleep_hours for each group

ggplot(data = my_data_clean, # starting data frame
       aes(x = work_night, # x axis as type of night
           y = total_sleep_hours)) + # y axis as total sleep hours 
  stat_summary(fun = mean, # show bar representing the mean
                geom = "bar",
                fill = "plum",
                width = 0.6) + 
  geom_jitter(width = 0.2, alpha = 0.6, color = "black", size = 1.5) + # show individual data points
  labs(x = "Work Night", y = "Total Sleep (hours)",
       title = "Comparison of Total Sleep Between Work and Non-Work Nights") + # relabel axes
  theme_minimal() # apply a clean theme 
```

## c. 

## d. 

# Problem 2. Affective Visualization

## a. 

## b. 

## c. 

## d. 

# Problem 3. Statistical Critique

## a.

## b. 

## c. 
